Are environmental and geographic variables effective surrogates for genetic variation in conservation planning?

### 2016-03-17 article/article.Rmd
+ update surrogate-based targets + increase percentages + replace drop1 with manual term removal + use minkowski distances instead of euclid + assorted fixes in manuscript

To understand the relationship between the surrogate-based targets used to generate a prioritization and the proportion of genetic variation it secures, we generated a collection of single-species prioritizations. For each species, we first generated two sets of prioritizations using increasing environmental and geographic-based targets (separately). Prioritizations were generated using surrogate-based targets ranging from `r (min(rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.target)*100)` % and `r (max(rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.target)*100)` % in increments of `r diff(rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.target[1:2])*100` % ($n = `r rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.optimal.replicates`$ per target). To determine if these surrogate-based prioritizations performed better then random, we then generated another two sets of prioritizations by randomly selecting planning units that were occupied by the species ($n = `r rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.random.replicates`$ per surrogate target). To account for variation in the number of selected planning units in the target-based prioritizations, the number of selected of planning units in the randomly generated prioritizations were generated following their distribution. Randomly generated prioritizations associated with negative values--because they secured only a very small amount of genetic variation--were replaced with zeros to facilitate statistical analysis.

@@

## Statistical analysis
We investigated the effectiveness of prioritizations generated using surrogate-based targets using generalized linear mixed-effects models [GLMMs; using a logit link and the $\texttt{lme4}$ R package; @r469]. To determine  if the environmental and geographic variables were effective surrogates, we fit a full model to the proportion of genetic variation that the first collection of prioritizations secured using categorical variables to indicate the surrogate-based targets used to generate them, the type of genetic variation (adaptive or neutral), the method using the generate the prioritization (optimal or random), and the interactions between these variables. To determine if surrogate-based targets improved the effectiveness of prioritizations under various scenarios, we fit a full model to the proportion of genetic variation that second collection of prioritizations secured using categorical variables to indicate the scenario used to generate the prioritization (single-species, multi-species, multi-species with cost), the combination of targets used to generate the prioritizations (amount-based targtes, amount and surrogate-based targets, amount and genetic-based targets), the type of genetic variation measured (adaptive or neutral), and the interactions between them. All model contained intercept- and slope-level random effects to accommodate variation between species. Both full models were subject to a step-wise term deletion routine using $\chi^{2}$ tests to assess significance (terms were retained when $P < 0.05$). Finally, we conducted post-hoc analysis on each full model using Tukey contrasts with a Bonferroni correction [using the $\texttt{multcomp}$ R package; @r470].

@@

### Limitations/Future directions
We wish to alert practitioners to several limitations associated with our analysis. First, the size of the planning units we used (23 $\times$ 25 km) is much larger than typically used in conservation planning scenarios (approx. 1--5 km^2). We used this resolution because the genetic data was collected at this scale. Whilst we could have interpolated the genetic data to a finer resolution and used smaller planning units, this would have biased our analysis because the surrogate variables are spatially auto-correlated. Second, we used geographic distances as surrogates for connectivity between planning units. Although connectivity between areas can be modeled using isolation by resistance measures [@r473] derived from landscape data [eg. @r474], they are unlikely to be of use in broad-scale conservation exercises since they need to be tuned using species-specific parameters. Third, we used AFLP data to characterize the main gradients of genetic variation. Whilst next-generation sequencing provides higher resolution genetic information than AFLP data [reviewed in @r480], our methodology would still have utilized only the main gradients of the genetic variation to generate prioritizations in a feasible amount of time. Overall, we are confident that the limitations associated with our study are relatively minor.

>>>>

To understand the relationship between the surrogate-based targets used to generate a prioritization and the proportion of genetic variation it secures, we generated a collection of single-species prioritizations. For each species, we first generated two sets of prioritizations using increasing environmental and geographic-based targets (separately). Prioritizations were generated using surrogate-based targets ranging from `r (min(rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.target)*100)` % and `r (max(rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.target)*100)` % in increments of `r diff(rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.target[1:2])*100` % ($n = `r rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.optimal.replicates`$ per target). To determine if these surrogate-based prioritizations performed better then random, we then generated another two sets of prioritizations by randomly selecting planning units that were occupied by the species ($n = `r rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.random.replicates`$ per surrogate target). To account for variation in the number of selected planning units in the target-based prioritizations, the number of selected planning units in the randomly generated prioritizations was equal to the target-based prioritizations. Randomly generated prioritizations associated with negative values--because they secured only a very small amount of genetic variation--were replaced with zeros to facilitate statistical analysis.

@@

## Statistical analysis
We investigated the effectiveness of prioritizations generated using surrogate-based targets using generalized linear mixed-effects models [GLMMs; using a logit link and the $\texttt{lme4}$ R package; @r469]. To determine  if the environmental and geographic variables were effective surrogates, we fit a full model to the proportion of genetic variation that the first collection of prioritizations secured using categorical variables to indicate the surrogate-based targets used to generate them, the type of genetic variation (adaptive or neutral), the method used to generate the prioritization (optimal or random), and the interactions between these variables. To determine if surrogate-based targets improved the effectiveness of prioritizations under various scenarios, we fit a full model to the proportion of genetic variation that second collection of prioritizations secured using categorical variables to indicate the scenario (single-species, multi-species, multi-species with cost), the combination of targets used to generate the prioritizations (amount-based targets, amount and surrogate-based targets, amount and genetic-based targets), the type of genetic variation measured (adaptive or neutral), and the interactions between them. All model contained intercept-level random effects to accommodate variation between species. Both full models were subject to a step-wise term deletion routine using $\chi^{2}$ tests to assess significance (terms were retained when $P < 0.05$). Finally, we conducted post-hoc analysis on each full model using Tukey contrasts with a Bonferroni correction [using the $\texttt{multcomp}$ R package; @r470].

@@

### Limitations/Future directions
We wish to alert our readers to several limitations associated with our analysis. First, the size of the planning units we used (23 $\times$ 25 km) is much larger than typically used in conservation planning scenarios (approx. 1--5 km^2). We used this resolution because the genetic data was collected at this scale. Whilst we could have interpolated the genetic data to a finer resolution and used smaller planning units, this would have biased our analysis because this process would have introduced artificial spatial auto-correlation into the dataset. Second, we used geographic distances as surrogates for connectivity between planning units. Although connectivity between areas can be modeled using isolation by resistance measures [@r473] derived from landscape data [eg. @r474], they are unlikely to be of use in broad-scale conservation exercises since they need to be tuned using species-specific parameters. Third, we used AFLP data to characterize the main gradients of genetic variation. Whilst next-generation sequencing provides higher resolution genetic information than AFLP data [reviewed in @r480], our methodology would still have utilized only the main gradients of the genetic variation in order to generate prioritizations in a feasible amount of time. Overall, we are confident that the limitations associated with our study are relatively minor.
